---
title: "PA1_template"
output:
  html_document:
    pandoc_args: [
      "+RTS", "-K64m",
      "-RTS"
    ]
---

Reproducible Research Exam - Peer Assesment n.1

###STEP 1 - Loading and preprocessing the data  
  
Loading Libraries
```{r}
library(plyr)
library(dplyr)
```

  
Loading and Analyzing data
```{r}

db<-read.csv("activity.csv")

head(db)
str(db)
summary(db)
```


###STEP 2 - What is mean total number of steps taken per day?


1. Calculate the total number of steps taken per day

```{r}
db_mean_day<-group_by(db,date)
sums_by_day<-summarize(db_mean_day,sum(steps,na.rm=TRUE))
colnames(sums_by_day)[2]<-"sum_Steps"
head(sums_by_day)
```

2. If you do not understand the difference between a histogram and a barplot, research the difference between them. Make a histogram of the total number of steps taken each day

```{r}
hist(sums_by_day$sum_Steps)
```


3. Calculate and report the mean and median of the total number of steps taken per day

```{r}
means_by_day<-summarize(db_mean_day,mean(steps,na.rm=TRUE))
colnames(means_by_day)[2]<-"mean_Steps"
means_by_day$mean_Steps[is.nan(means_by_day$mean_Steps)]<-0
head(means_by_day)

median_by_day<-summarize(db_mean_day,median(steps,na.rm=TRUE))
colnames(median_by_day)[2]<-"median_Steps"
median_by_day$median_Steps[is.nan(median_by_day$median_Steps)]<-0
head(median_by_day)
```



###STEP 3 -What is the average daily activity pattern?

1. Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r}
db_mean_interval<-group_by(db,interval)
means_by_interval<-summarize(db_mean_interval,mean(steps,na.rm=TRUE))
colnames(means_by_interval)[2]<-"mean_Steps"
plot(means_by_interval$interval, means_by_interval$mean_Steps, type = "l", col = "black", ylab = "mean n. steps", xlab = "intervals")
```

2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
max_interval<-subset(means_by_interval,means_by_interval$mean_Steps==max(means_by_interval$mean_Steps))
max_interval
```


###STEP 4 - Imputing missing values

1. Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)

```{r}
any(is.na(db))

db_na_logic=is.na(db)
nrow(subset(db,db_na_logic))
```

2. Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.

3. Create a new dataset that is equal to the original dataset but with the missing data filled in.
   **I choose to replace missing values with the mean of that day**
```{r}
db_with_mean<-join(db,means_by_day,type="left")

for(i in 1:nrow(db_with_mean)){
  if (is.na(db_with_mean[i,1])) {
    db_with_mean[i,1]<- db_with_mean[i,4]
  } else {    
  }}

db_with_mean<-db_with_mean[,c(1,2,3)]
head(db_with_mean)
```


4. Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. 

```{r}
db_mean_day_2<-group_by(db_with_mean,date)
sums_by_day_2<-summarize(db_mean_day_2,sum(steps,na.rm=TRUE))
colnames(sums_by_day_2)[2]<-"sum_Steps"
hist(sums_by_day_2$sum_Steps)

means_by_day_2<-summarize(db_mean_day_2,mean(steps,na.rm=TRUE))
colnames(means_by_day_2)[2]<-"mean_Steps"
means_by_day_2$mean_Steps[is.nan(means_by_day_2$mean_Steps)]<-0
head(means_by_day_2)

median_by_day_2<-summarize(db_mean_day_2,median(steps,na.rm=TRUE))
colnames(median_by_day_2)[2]<-"median_Steps"
head(median_by_day_2) 
```

Do these values differ from the estimates from the first part of the assignment? 
*NO that much, most NA values are in days without steps*

What is the impact of imputing missing data on the estimates of the total daily number of steps?
*Ther eis a very low impact, most replacement are 0s instead of NA*


###STEP 5 - Are there differences in activity patterns between weekdays and weekends?

1. Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day. --> *Sorry, weekdays are in Italian :)*

```{r}
db_with_mean$day<-weekdays(as.Date(db_with_mean$date))

for(i in 1:nrow(db_with_mean)){
  if (db_with_mean[i,4]=='Sabato' || db_with_mean[i,4]=='Domenica') {
    db_with_mean[i,4]="weekend"
  } else { 
    db_with_mean[i,4]="weekday"
  }}

head(db_with_mean)
```

2. Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). 

```{r}
db_with_mean_weekday<-subset(db_with_mean,day=="weekday")
db_with_mean_weekend<-subset(db_with_mean,day=="weekend")

db_mean_interval_weekday<-group_by(db_with_mean_weekday,interval)
means_by_interval_weekday<-summarize(db_mean_interval_weekday,mean(steps,na.rm=TRUE))
colnames(means_by_interval_weekday)[2]<-"mean_Steps"

db_mean_interval_weekend<-group_by(db_with_mean_weekend,interval)
means_by_interval_weekend<-summarize(db_mean_interval_weekend,mean(steps,na.rm=TRUE))
colnames(means_by_interval_weekend)[2]<-"mean_Steps"


par(mfrow=c(1,2))
plot(means_by_interval_weekday$interval, means_by_interval_weekday$mean_Steps, type = "l", col = "black", ylab = "mean n. steps weekday", xlab = "intervals")
plot(means_by_interval_weekend$interval, means_by_interval_weekend$mean_Steps, type = "l", col = "red", ylab = "mean n. steps weekend", xlab = "intervals")
```
